---
title: "Roe Deer Logistic Regression Analysis"
author: "Ashlee Mikkelsen"
date: "`r Sys.Date()`"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Introduction

This is the logistic regression file for the Roe deer- vehicle collision analysis for Viken kommune. Andreas, Rick and I took a contract from the kommune to estimate whether decreasing the speed limit on roads by 10 km/hr would reduce the number of vehicle collisions with Roe deer. 

# Environment

```{r}

rm(list=ls())

```

## Set working directory

```{r}

setwd("~Rprojects\RoeDeerVehCol")

```


## Load packages

```{r}

library(ggplot2)
library(lme4)
library(wiqid)
library(AICcmodavg)
library(ggeffects)


```


## set graphing theme

```{r, include=FALSE}

mytheme <- theme(
    axis.text = element_text(size = 12,face = "bold"),
    axis.title = element_text(size = 14, face = "bold"),
    panel.grid.major = element_line(color = "grey92"),
    panel.grid.minor = element_line(color = "grey96"),
    panel.background = element_rect(fill = "white"),
    axis.line = element_line(colour = "black",linewidth = 1),
    axis.ticks = element_line(linewidth = 1),
    legend.position = "top"
    )
theme_set(mytheme)

P <- palette(viridis(40))
P <- palette(viridis(40))

```


## Functions

```{r}

logit2prob <- function(logit){
  odds <- exp(logit)
  prob <- odds / (1 + odds)
  return(prob)
}



rsq <- function (x, y) cor(x, y) ^ 2

```


## Load Data

```{r}

RoeDeer.df <- read.csv("Data\\df_VikenRoeDeer_7080Combined.csv",
                         header=TRUE)
RoeDeer.df$f.speed <- as.factor(RoeDeer.df$speed)

RoeDeer.df2 <- RoeDeer.df


```

The original data file has 21090 observations and was processed by Rick prior to analysis. Explanation of data file:
Columns
1: Object ID- Unique identifyer for each record
2: id- secondary identifyer identifies the response category of the observation and the observation number within the response category (1_1 is the first observation of the 1 response category)
3: response- category 1= a collision 0= randomly drawn background points that are psuedo-absence points.
4: X- continuous variable UTM northing of the record
5: Y- continuous variable UTM easting of the record
6: speed- discrete numeric Speed limit of road segment closest straight line distance from the record
7: ADT- discrete numeric Average daily traffic volume of road segment closest straight-line distance from the record.
8: efficiency-  continuous variable (0-1) indexing the straightness of the road in which 0= perfectly straigh and records greater than 0 have some level curvature. 
9: cover_fo- continuous variable (0-1) proportion of forest cover calculated as
10: cover_agri- continuous variable (0-1) proportion of forest cover calculated as
11: cover_wa- continuous variable (0-1) proportion of water calculated as
12: cover_urban- continuous variable (0-1) proportion of urban land cover calculated as
13: dist_cover- continuous variable staraight line distance from the record to (forest?) cover, calculated as, units= meters?
14: dist-agri- continuous variable staraight line distance from the record to agriculture land cover, calculated as, units= meters?
15: dist_water- continuous variable staraight line distance from the record to (forest?) cover, calculated as, units= meters?
16: dist_urban- continuous variable staraight line distance from the record to (forest?) cover, calculated as, units= meters?
17: geometry-
18: f.speed- categorical variable for speed


## Standardize continuos variables

```{r}



```


# Models

## Intercept only

```{r}
Models<-list()

Models[[1]]<-INT<-glm(response~1,
                        data = RoeDeer.df,
                        family = binomial(link="logit"))
summary(INT)

Prob.Int <- logit2prob(INT$coefficients[1])

```


## Speed

### continuous discrete

```{r}

Models[[2]]<-SPD.c<-glm(response ~ speed,
                        data = RoeDeer.df,
                        family = binomial(link="logit"))
summary(SPD.c)

spd.c_coef <- logit2prob(SPD.c$coefficients[2])

```

### categorical (factor)

```{r}

Models[[3]]<-SPD.f<-glm(response ~ f.speed,
                        data = RoeDeer.df,
                        family = binomial(link="logit"))
summary(SPD.f)


```


### Traffic volume
```{r}
Models[[4]]<-SPD.c_Traf<-glm(response ~ speed+ADT,
                        data = RoeDeer.df,
                        family = binomial(link="logit"))
```

#### Model output

```{r}

summary(SPD.c_Traf)


Prob.40 <- logit2prob(SPD.c_Traf$coefficients[1]+
  (SPD.c_Traf$coefficients[2]*40)+
  (SPD.c_Traf$coefficients[3]*mean(RoeDeer.df$ADT)))


Prob.50 <- logit2prob(SPD.c_Traf$coefficients[1]+
  (SPD.c_Traf$coefficients[2]*50)+
  (SPD.c_Traf$coefficients[3]*mean(RoeDeer.df$ADT)))

Prob.60 <- logit2prob(SPD.c_Traf$coefficients[1]+
  (SPD.c_Traf$coefficients[2]*60)+
  (SPD.c_Traf$coefficients[3]*mean(RoeDeer.df$ADT)))

Prob.75 <- logit2prob(SPD.c_Traf$coefficients[1]+
  (SPD.c_Traf$coefficients[2]*75)+
  (SPD.c_Traf$coefficients[3]*mean(RoeDeer.df$ADT)))



RoeDeer.df$FIT<-fitted(SPD.c_Traf)

PredColl<-ggpredict(SPD.c_Traf, terms = "speed")

ggplot()+
  geom_errorbar(data = PredColl,
                aes(x, predicted,ymin=conf.low, ymax=conf.high),
                width=0.3, linewidth=1)+
  geom_point(data = PredColl, aes(x, predicted),
             size=2,
             color="red")+
  scale_x_continuous(expand = c(0.01,0.01))+
  scale_y_continuous(expand = c(0,0),
                     breaks = seq(0,1,0.1),
                     limits = c(0,0.5))+
  ylab("Probability of collision")+
  xlab("Speed limit (km/h)")
 

```


# Generate model table for comparison

```{r}

Modnames <- c("INTERCEPT", "SPEEDc", "SPEEDf", "SPEEDc_TRAFFIC")

##generate AICc table

aictab(cand.set = Models,
       modnames = Modnames, sort = TRUE, second.ord=TRUE)

##round to 4 digits after decimal point and give log-likelihood
print(aictab(cand.set = Models,
             modnames = Modnames,
             sort = TRUE),
      digits = 4,
      LL = TRUE)

write.csv(aictab(cand.set = Models,
                 modnames = Modnames,
                 sort = TRUE,
                 second.ord=TRUE),
          file="~Results\\ModelSelectionTable.csv")


```

