---
title: "Roe Deer Logistic Regression Analysis"
author: "Ashlee Mikkelsen"
date: "`r Sys.Date()`"
output: word_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE)

```


# Introduction

This is the logistic regression file for the Roe deer- vehicle collision analysis for Viken kommune. Andreas, Rick and I took a contract from the kommune to estimate whether decreasing the speed limit on roads by 10 km/hr would reduce the number of vehicle collisions with Roe deer. 

# Environment

```{r}

rm(list=ls())

```

## Set working directory

```{r}

setwd("~Rprojects\\RoeDeerVehCol")

```


## Load packages

```{r}

library(ggplot2)
library(lme4)
library(wiqid)
library(AICcmodavg)
library(ggeffects)
library(corrplot)
library(viridis)


```


## set graphing theme

```{r, include=FALSE}

mytheme <- theme(
    axis.text = element_text(size = 12,face = "bold"),
    axis.title = element_text(size = 14, face = "bold"),
    panel.grid.major = element_line(color = "grey92"),
    panel.grid.minor = element_line(color = "grey96"),
    panel.background = element_rect(fill = "white"),
    axis.line = element_line(colour = "black",linewidth = 1),
    axis.ticks = element_line(linewidth = 1),
    legend.position = "top"
    )
theme_set(mytheme)


```


## Functions

```{r}

logit2prob <- function(logit){
  odds <- exp(logit)
  prob <- odds / (1 + odds)
  return(prob)
}



rsq <- function (x, y) cor(x, y) ^ 2

```


## Load Data

```{r}

RoeDeer.df <- read.csv("Data\\df_VikenRoeDeer_7080Combined.csv",
                         header=TRUE)
RoeDeer.df$f.speed <- as.factor(RoeDeer.df$speed)
RoeDeer.df$f.speed_comb <- as.factor(RoeDeer.df$speed_comb)


```


### Data description

The original data file has 21090 observations and was processed by Rick prior to analysis. Explanation of data file:
Columns
1: Object ID- Unique identifyer for each record
2: id- secondary identifyer identifies the response category of the observation and the observation number within the response category (1_1 is the first observation of the 1 response category)
3: response- category 1= a collision 0= randomly drawn background points that are psuedo-absence points.
4: X- continuous variable UTM northing of the record in ETRS 98 / UTM 33N (epsg code: 25833).
5: Y- continuous variable UTM easting of the record in ETRS 98 / UTM 33N (epsg code: 25833).
6: speed- discrete numeric Speed limit of road segment closest straight line distance from the record
7: ADT- discrete numeric Average daily traffic volume of road segment closest straight-line distance from the record.
8: efficiency-  continuous variable (0-1) indexing the straightness of the road in which 1= perfectly straight and records less than 1 have some level of curvature. 
9: cover_fo- continuous variable (0-1) proportion of forest cover calculated as area of habitat variable within a 100m buffer divided by the area of the 100m buffer * 100. 
10: cover_agri- continuous variable (0-1) proportion of forest cover calculated as area of habitat variable within a 100m buffer divided by the area of the 100m buffer * 100. 
11: cover_wa- continuous variable (0-1) proportion of water calculated as area of habitat variable within a 100m buffer divided by the area of the 100m buffer * 100. 
12: cover_urban- continuous variable (0-1) proportion of urban land cover calculated as area of habitat variable within a 100m buffer divided by the area of the 100m buffer * 100. 
13: dist_cover- continuous variable straight line distance from the record to closest polygon where cover="FOREST", calculated as, units= meters?
14: dist-agri- continuous variable straight line distance from the record to agriculture land cover, calculated as, units= meters?
15: dist_water- continuous variable straight line distance from the record to (forest?) cover, calculated as, units= meters?
16: dist_urban- continuous variable straight line distance from the record to (forest?) cover, calculated as, units= meters?
17: geometry- Remnant variable from sf package. ignore
18: f.speed- categorical variable for speed


## Standardize continuos variables

```{r}

RoeDeer.df$zADT <- standardize(RoeDeer.df$ADT)
RoeDeer.df$zEFF <- standardize(RoeDeer.df$efficiency)
RoeDeer.df$zFOR <- standardize(RoeDeer.df$cover_fo)
RoeDeer.df$zAGR <- standardize(RoeDeer.df$cover_agri)
RoeDeer.df$zWAT <- standardize(RoeDeer.df$cover_wa)
RoeDeer.df$zURB <- standardize(RoeDeer.df$cover_urb)
RoeDeer.df$zDFR <- standardize(RoeDeer.df$dist_cover)
RoeDeer.df$zDAG <- standardize(RoeDeer.df$dist_agri)
RoeDeer.df$zDWA <- standardize(RoeDeer.df$dist_water)
RoeDeer.df$zDUR <- standardize(RoeDeer.df$dist_urban)


```

# Corrrelations

Below I look at the correlations between continuous variables in the dataset to determine which can be included together in the same model

```{r}


RD_NumCov <- RoeDeer.df[,c(6,8:16)]
C <- cor(RD_NumCov)


corrplot(C,
         type = "upper",
         order="hclust",
         addCoef.col = "black")


```

There are a few covariates that are highly correlated and cannot be included in the same model:
The amount of urban cover is highly correlated (-0.64) with speed limit
Forest cover is highly correlated with distance to forest (-0.63) and distance to agriculture (-0.59).

# Univariate Models

## Intercept only

```{r}
UnivariateModels<-list()

UnivariateModels[[1]]<-INT<-glm(response~1,
                        data = RoeDeer.df,
                        family = binomial(link="logit"))
summary(INT)

Prob.Int <- logit2prob(INT$coefficients[1])

```


## Speed

### continuous discrete

```{r}

UnivariateModels[[2]]<-SPD.c<-glm(response ~ speed,
                        data = RoeDeer.df,
                        family = binomial(link="logit"))


```

### categorical (factor)

```{r}

UnivariateModels[[3]]<-SPD.f<-glm(response ~ f.speed_comb,
                        data = RoeDeer.df,
                        family = binomial(link="logit"))


```


## Traffic volume
```{r}

UnivariateModels[[4]]<-TRAF<-glm(response ~ zADT,
                        data = RoeDeer.df,
                        family = binomial(link="logit"))
```


## Forest Cover

```{r}

UnivariateModels[[5]]<-FORCOV<-glm(response ~ zFOR,
                        data = RoeDeer.df,
                        family = binomial(link="logit"))

```


## Agriculture Cover

```{r}

UnivariateModels[[6]]<-AGCOV<-glm(response ~ zAGR,
                        data = RoeDeer.df,
                        family = binomial(link="logit"))

```


## Water Cover

```{r}

UnivariateModels[[7]]<-WACOV<-glm(response ~ zWAT,
                        data = RoeDeer.df,
                        family = binomial(link="logit"))

```


## Urban cover

```{r}

UnivariateModels[[8]]<-URBCOV<-glm(response ~ zURB,
                        data = RoeDeer.df,
                        family = binomial(link="logit"))

```


## Distance to forest

```{r}

UnivariateModels[[9]]<-DISFOR<-glm(response ~ zDFR,
                        data = RoeDeer.df,
                        family = binomial(link="logit"))

```


## Distance to Agriculture

```{r}

UnivariateModels[[10]]<-DISTAG<-glm(response ~ zDAG,
                        data = RoeDeer.df,
                        family = binomial(link="logit"))

```


## Distance to water

```{r}

UnivariateModels[[11]]<-DISWA<-glm(response ~ zDWA,
                        data = RoeDeer.df,
                        family = binomial(link="logit"))

```


## Road Efficiency

```{r}

UnivariateModels[[12]]<-EFF<-glm(response ~ zEFF,
                        data = RoeDeer.df,
                        family = binomial(link="logit"))
```


## Distance to urban

```{r}

UnivariateModels[[13]]<-DUR<-glm(response ~ zDUR,
                        data = RoeDeer.df,
                        family = binomial(link="logit"))
```




## Univariate AICc table

```{r}

UnivariateModnames <- c("INTERCEPT", "SPEEDc", "SPEEDf", "TRAFFIC",
                        "FOREST COVER", "AG COVER", "WATER COVER","URBAN COVER",
                        "DIST FOREST","DIST AG","DIST WATER", "EFFICIENCY",
                        "DIST URBAN")


##Generate and print AICc table
print(aictab(cand.set = UnivariateModels,
             modnames = UnivariateModnames,
             sort = TRUE),
      digits = 4, #round to 4 digits after decimal point
      LL = TRUE #give log-likelihood 
      )

write.csv(aictab(cand.set = UnivariateModels,
                 modnames = UnivariateModnames,
                 sort = TRUE,
                 second.ord=TRUE),
          file="Results\\UnivariateModelSelectionTable.csv")


```


Based on the AICc table, all of the covariates did a better job at explaining variation in the data than the intercept only, so I will carry all covariates through

# Bivariate models with focus on speed

Distance to forest was the best univariate model, so I will include it in the bivariate model list as a comparison.

The categorical covariate for speed out-performed the continuous covariate, so I will use that one in further modeling steps.

```{r}

BivariateModels<-list()

BivariateModels[[1]]<-DISFOR

```


## Speed & Distance to forest

```{r}

BivariateModels[[2]]<-SPD.f_DFOR<-glm(response ~ f.speed_comb+zDFR,
                        data = RoeDeer.df,
                        family = binomial(link="logit"))

```




## Urban cover & Distance to forest

```{r}

BivariateModels[[3]]<-URBC_DFOR<-glm(response ~ zURB+zDFR,
                        data = RoeDeer.df,
                        family = binomial(link="logit"))

```



## Speed & Forest Cover

```{r}

BivariateModels[[4]]<-SPD.f_FORC<-glm(response ~ f.speed_comb+zFOR,
                        data = RoeDeer.df,
                        family = binomial(link="logit"))

```


## Speed & Efficiency

```{r}

BivariateModels[[5]]<-SPD.f_EFF<-glm(response ~ f.speed_comb+zEFF,
                        data = RoeDeer.df,
                        family = binomial(link="logit"))

```



## Speed & Traffic

```{r}

BivariateModels[[6]]<-SPD.f_ADT<-glm(response ~ f.speed_comb+zADT,
                        data = RoeDeer.df,
                        family = binomial(link="logit"))

```



## Full speed model (no interactions)

Urban cover excluded because it is highly correlated with speed and distance to forest is included rather than forest cover because it performed better than forest cover

```{r}

BivariateModels[[7]]<-SPD.f_FULL<-glm(response ~ f.speed_comb+
                                        zDFR+
                                        zEFF+
                                        zADT+
                                        zWAT,
                        data = RoeDeer.df,
                        family = binomial(link="logit"))

```



## Full urban cover model (no interactions)

Urban cover excluded because it is highly correlated with speed


```{r}

BivariateModels[[8]]<-URB_FULL<-glm(response ~ zURB+
                                        zDFR+
                                        zEFF+
                                        zADT+
                                        zWAT,
                        data = RoeDeer.df,
                        family = binomial(link="logit"))

```


## Full speed model (SPD*DFR)

Urban cover excluded because it is highly correlated with speed and distance to forest is included rather than forest cover because it performed better than forest cover

```{r}

BivariateModels[[9]]<-SPD.f_FULL_INX<-glm(response ~ f.speed_comb+
                                        zDFR+
                                        zEFF+
                                        zADT+
                                        zWAT+
                                        f.speed*zDFR,
                        data = RoeDeer.df,
                        family = binomial(link="logit"))

```




## Bivariate AICc table

```{r}

BivariateModnames <- c("DIST FOREST","SPEEDc_DIST FOREST",
                       "URBAN COVER_DIST FOR", "SPEEDf_FOREST COVER",
                       "SPEEDf_EFFICIENCY", "SPEEDf_TRAFFIC",
                       "SPEEDf_FULL", "URBAN COVER_FULL", "SPEEDf_FULL_INX")


## generate and print AICc table
print(aictab(cand.set = BivariateModels,
             modnames = BivariateModnames,
             sort = TRUE),
      digits = 4, #round to 4 digits after decimal point
      LL = TRUE #give log-likelihood
      )

write.csv(aictab(cand.set = BivariateModels,
                 modnames = BivariateModnames,
                 sort = TRUE,
                 second.ord=TRUE),
          file="Results\\BivariateModelSelectionTable.csv")


```



# Speed only output

## Numerical

```{r}

summary(SPD.c)

dummy.speed <- seq(min(RoeDeer.df$speed),
                   max(RoeDeer.df$speed),
                   length=100)

SpeedEqn <- function(x){
  SPD.c$coefficients[1]+
    SPD.c$coefficients[2]*x
}

PredLogit.speed <- SpeedEqn(dummy.speed)
PredProb.Speed <- logit2prob(PredLogit.speed)

Prob10 <- logit2prob(SpeedEqn(60))-logit2prob(SpeedEqn(50))

Speed.df <- as.data.frame(cbind(dummy.speed,
                                PredProb.Speed))

Speed.Fvalues<-fitted(SPD.c)
Speed.eFIT<-exp(fitted(SPD.c))
RoeDeer.df$Speed.FIT<-Speed.Fvalues
RoeDeer.df$Speed.eFIT<-Speed.eFIT


SPD<-ggpredict(SPD.c, terms = "speed")
print(SPD)

plot(SPD, residuals =TRUE)



```



## Speed Graph

```{r}

SpdPlot <- ggplot(data = SPD, aes(x, predicted))+
  geom_ribbon(aes(ymin=conf.low, ymax=conf.high), alpha=0.1)+
  geom_line(lwd=1.1)+
  scale_x_continuous(expand=c(0,0))+
  scale_y_continuous(expand = c(0,0),
                     limits = c(0,0.5),
                     breaks = seq(0,0.5,0.1))+
  ylab("Probability of roe deer - vehicle collision")+
  xlab("Nearest posted speed limit (km/h)")+
  geom_text(label="Increase of 10 km/h =\n~ 4% increase in collision probability",
            size=4,
            x=50,y=0.35)+
  geom_segment(aes(x=50, y=0.255,
                   xend=60, yend=0.255),
               linewidth=1.2,
               linetype="dashed",
               color="red4")+
  geom_segment(aes(x=60, y=0.255,
                   xend=60, yend=0.295),
               linewidth=1.2,
               linetype="dashed",
               color="red4")
  

SpdPlot

ggsave(plot = SpdPlot,
       "Spd4Reduc.png",
       height=5, width = 6,
       units = "in",
       dpi=700)

```



# Top Model output (Urban cover)

## Numerical output

```{r}

summary(URB_FULL)

#### Distance to forest

zDummy.ForDist <- seq(min(RoeDeer.df$zDFR),
                     max(RoeDeer.df$zDFR),
                     length=100)

ForDist.TM <- function(x){
  URB_FULL$coefficients[1]+
    URB_FULL$coefficients[2]*mean(RoeDeer.df$zURB)+
    URB_FULL$coefficients[3]*x+
    URB_FULL$coefficients[4]*mean(RoeDeer.df$zEFF)+
    URB_FULL$coefficients[5]*mean(RoeDeer.df$zADT)+
    URB_FULL$coefficients[6]*mean(RoeDeer.df$zWAT)+
    URB_FULL$coefficients[7]*mean(RoeDeer.df$zDAG)
}

zPredlogit.DistFor <- ForDist.TM(zDummy.ForDist)
zPredProb.DistFor <- logit2prob(zPredlogit.DistFor)

dummy.ForDist <- (zDummy.ForDist*sd(RoeDeer.df$dist_cover))+
  mean(RoeDeer.df$dist_cover)

log(0.04/(1-0.04))
(((-3.18+0.75)/-0.33)*sd(RoeDeer.df$dist_cover))+mean(RoeDeer.df$dist_cover)

ggplot(data = )

##### Urban cover

zDummy.UrbanCover <- seq(min(RoeDeer.df$zURB),
                     max(RoeDeer.df$zURB),
                     length=100)

UrbCov.TM <- function(x){
  URB_FULL$coefficients[1]+
    URB_FULL$coefficients[2]*x+
    URB_FULL$coefficients[3]*mean(RoeDeer.df$zDFR)+
    URB_FULL$coefficients[4]*mean(RoeDeer.df$zEFF)+
    URB_FULL$coefficients[5]*mean(RoeDeer.df$zADT)+
    URB_FULL$coefficients[6]*mean(RoeDeer.df$zWAT)+
    URB_FULL$coefficients[7]*mean(RoeDeer.df$zDAG)
}

zPredlogit.UrbCov <- UrbCov.TM(zDummy.UrbanCover)
zPredProb.UrbCov <- logit2prob(zPredlogit.UrbCov)

dummy.UrbCover <- (zDummy.UrbanCover*sd(RoeDeer.df$cover_urb))+
  mean(RoeDeer.df$cover_urb)



###### Data frame

URB_FULL.df <- as.data.frame(cbind(dummy.ForDist,
                                   zPredProb.DistFor,
                                   dummy.UrbCover,
                                   zPredProb.UrbCov))


UrbanFull_DF.Fvalues<-fitted(URB_FULL)
UrbanFull_DF.eFIT<-exp(fitted(URB_FULL))
RoeDeer.df$UrbanFull_DF.FIT<-UrbanFull_DF.Fvalues
RoeDeer.df$UrbanFull_DF.eFIT<-UrbanFull_DF.eFIT


UrbanFull_DF<-ggpredict(URB_FULL, terms = "zDFR[all]")
print(UrbanFull_DF)
UrbanFull_DF$aX <- (UrbanFull_DF$x*sd(RoeDeer.df$dist_cover))+
  mean(RoeDeer.df$dist_cover)
plot(UrbanFull_DF, residuals =TRUE)


```


## Graphical output

```{r}

####### Distance to Forest

logit2prob(-0.75)+logit2prob(-0.33)

ggplot(data = URB_FULL.df, aes(dummy.ForDist,
                               zPredProb.DistFor))+
  geom_line()+
  geom_point()+
  scale_x_continuous(expand = c(0,0))+
  scale_y_continuous(expand = c(0,0),
                     limits = c(0,0.5),
                     breaks = seq(0,0.5,0.1))+
  ylab("Probability of roe deer - vehicle collision")+
  xlab("Distance to forest cover (meters)")
####################




Dist4Reduc <- ggplot(data = UrbanFull_DF, aes(aX, predicted))+
  geom_ribbon(aes(ymin=conf.low, ymax=conf.high), alpha=0.1)+
  geom_line(lwd=1.1)+
  scale_x_continuous(expand = c(0,0))+
  scale_y_continuous(expand = c(0,0),
                     limits = c(0,0.5),
                     breaks = seq(0,0.5,0.1))+
  ylab("Probability of  roe deer - vehicle collision")+
  xlab("Distance to forest cover (meters)")+
  geom_segment(aes(y=0.375, x=0,
                   yend=0.375, xend=67.5),
             linewidth=1.2,
             color="grey60",
             alpha=0.6,
             linetype="dashed")+
   geom_segment(aes(y=0.335, x=67.5,
                   yend=0.375, xend=67.5),
             linewidth=1.2,
             color="grey60",
             alpha=0.6,
             linetype="dashed")+
  geom_segment(aes(y=0.335, x=0,
                   yend=0.335, xend=67.5),
               linewidth=1.2,
               color="red4",
               alpha=0.6,
               linetype="dashed")+
  geom_segment(aes(y=0, x=67.5,
                   yend=0.335, xend=67.5),
               linewidth=1.2,
               color="red4",
               alpha=0.6,
               linetype="dashed")+
  geom_line(linewidth=1.1)+
  geom_text(label="~ 65 m increase in distance to forest corresponds\nto a 4% reduction in collision probability",
            x=350,y=0.35,
            size=3)+
  geom_text(label="65 m",
            size=4,
            x=105, y=0.01,
            color="red4")
Dist4Reduc

ggsave(plot = Dist4Reduc,
       "Dist4Reduc.png",
       height=5, width = 6,
       units = "in",
       dpi=700)

###### Urban Cover

ggplot(data =URB_FULL.df, aes(dummy.UrbCover,
       zPredProb.UrbCov))+
  geom_line(linewidth=1.1)+
  geom_point()+
  scale_x_continuous(expand=c(0,0))+
  scale_y_continuous(expand = c(0,0),
                     limits = c(0,0.5),
                     breaks = seq(0,0.5,0.1))+
  ylab("Probability of roe deer - vehicle collision")+
  xlab("Proportion of urban cover within 100 m buffer of collision")



```




# Secondary Model output (categorical speed)

## Numerical output

```{r}

summary(SPD.f_FULL)
mean(RoeDeer.df$speed)

#### Distance to forest

zDummy.ForDist <- seq(min(RoeDeer.df$zDFR),
                     max(RoeDeer.df$zDFR),
                     length=100)
## At 80 kmh
ForDist.Spd80.SM <- function(x){
  SPD.f_FULL$coefficients[1]+
    SPD.f_FULL$coefficients[2]*0+
    SPD.f_FULL$coefficients[3]*0+
    SPD.f_FULL$coefficients[4]*0+
    SPD.f_FULL$coefficients[5]*1+
    SPD.f_FULL$coefficients[6]*x+
    SPD.f_FULL$coefficients[7]*mean(RoeDeer.df$zEFF)+
    SPD.f_FULL$coefficients[8]*mean(RoeDeer.df$zADT)+
    SPD.f_FULL$coefficients[9]*mean(RoeDeer.df$zWAT)+
    SPD.f_FULL$coefficients[10]*mean(RoeDeer.df$zDAG)
}

zPredlogit.DistFor.Spd80 <- ForDist.Spd80.SM(zDummy.ForDist)
zPredProb.DistFor.Spd80 <- logit2prob(zPredlogit.DistFor.Spd80)

dummy.ForDist <- (zDummy.ForDist*sd(RoeDeer.df$dist_cover))+
  mean(RoeDeer.df$dist_cover)

## At 60

ForDist.Spd60.SM <- function(x){
  SPD.f_FULL$coefficients[1]+
    SPD.f_FULL$coefficients[2]*0+
    SPD.f_FULL$coefficients[3]*0+
    SPD.f_FULL$coefficients[4]*1+
    SPD.f_FULL$coefficients[5]*0+
    SPD.f_FULL$coefficients[6]*x+
    SPD.f_FULL$coefficients[7]*mean(RoeDeer.df$zEFF)+
    SPD.f_FULL$coefficients[8]*mean(RoeDeer.df$zADT)+
    SPD.f_FULL$coefficients[9]*mean(RoeDeer.df$zWAT)+
    SPD.f_FULL$coefficients[10]*mean(RoeDeer.df$zDAG)
}

zPredlogit.DistFor.Spd60 <- ForDist.Spd60.SM(zDummy.ForDist)
zPredProb.DistFor.Spd60 <- logit2prob(zPredlogit.DistFor.Spd60)

## At 50

ForDist.Spd50.SM <- function(x){
  SPD.f_FULL$coefficients[1]+
    SPD.f_FULL$coefficients[2]*0+
    SPD.f_FULL$coefficients[3]*1+
    SPD.f_FULL$coefficients[4]*0+
    SPD.f_FULL$coefficients[5]*0+
    SPD.f_FULL$coefficients[6]*x+
    SPD.f_FULL$coefficients[7]*mean(RoeDeer.df$zEFF)+
    SPD.f_FULL$coefficients[8]*mean(RoeDeer.df$zADT)+
    SPD.f_FULL$coefficients[9]*mean(RoeDeer.df$zWAT)+
    SPD.f_FULL$coefficients[10]*mean(RoeDeer.df$zDAG)
}

zPredlogit.DistFor.Spd50 <- ForDist.Spd50.SM(zDummy.ForDist)
zPredProb.DistFor.Spd50 <- logit2prob(zPredlogit.DistFor.Spd50)


## At 40

ForDist.Spd40.SM <- function(x){
  SPD.f_FULL$coefficients[1]+
    SPD.f_FULL$coefficients[2]*1+
    SPD.f_FULL$coefficients[3]*0+
    SPD.f_FULL$coefficients[4]*0+
    SPD.f_FULL$coefficients[5]*0+
    SPD.f_FULL$coefficients[6]*x+
    SPD.f_FULL$coefficients[7]*mean(RoeDeer.df$zEFF)+
    SPD.f_FULL$coefficients[8]*mean(RoeDeer.df$zADT)+
    SPD.f_FULL$coefficients[9]*mean(RoeDeer.df$zWAT)+
    SPD.f_FULL$coefficients[10]*mean(RoeDeer.df$zDAG)
}

zPredlogit.DistFor.Spd40 <- ForDist.Spd40.SM(zDummy.ForDist)
zPredProb.DistFor.Spd40 <- logit2prob(zPredlogit.DistFor.Spd40)



## At 30

ForDist.Spd30.SM <- function(x){
  SPD.f_FULL$coefficients[1]+
    SPD.f_FULL$coefficients[2]*0+
    SPD.f_FULL$coefficients[3]*0+
    SPD.f_FULL$coefficients[4]*0+
    SPD.f_FULL$coefficients[5]*0+
    SPD.f_FULL$coefficients[6]*x+
    SPD.f_FULL$coefficients[7]*mean(RoeDeer.df$zEFF)+
    SPD.f_FULL$coefficients[8]*mean(RoeDeer.df$zADT)+
    SPD.f_FULL$coefficients[9]*mean(RoeDeer.df$zWAT)+
    SPD.f_FULL$coefficients[10]*mean(RoeDeer.df$zDAG)
}

zPredlogit.DistFor.Spd30 <- ForDist.Spd30.SM(zDummy.ForDist)
zPredProb.DistFor.Spd30 <- logit2prob(zPredlogit.DistFor.Spd30)




###### Data frame

SPD_FULL.df <- as.data.frame(cbind(dummy.ForDist,
                                   zPredProb.DistFor.Spd80,
                                   zPredProb.DistFor.Spd60,
                                   zPredProb.DistFor.Spd50,
                                    zPredProb.DistFor.Spd40,
                                   zPredProb.DistFor.Spd30))

Testdf<-ggpredict(SPD.f_FULL, terms=c("zDFR [all]","f.speed_comb"))
Testdf$Xa <- (Testdf$x*sd(RoeDeer.df$dist_cover))+
  mean(RoeDeer.df$dist_cover)


```



## Graphical output


```{r}

Spd_DistFor <- ggplot(data = Testdf, aes(Xa, predicted))+
  geom_ribbon(aes(ymin=conf.low, ymax=conf.high,
                  fill=group), alpha=0.15)+
  geom_line(aes(color=group,
                lty=group),
            lwd=1.6)+
  scale_x_continuous(expand=c(0,0))+
  scale_y_continuous(expand = c(0,0),
                     limits = c(0,0.5),
                     breaks = seq(0,0.5,0.1))+
  ylab("Probability of roe deer - vehicle collision")+
  xlab("Distance to forest cover (meters)")+
  scale_color_manual(values = c("cadetblue1",
                                "turquoise1",
                                "turquoise2",
                                "turquoise3",
                                "turquoise4"),
                     labels=c("30","40","50","60","70/80"))+
  scale_fill_manual(values = c("cadetblue1",
                                "turquoise1",
                                "turquoise2",
                                "turquoise3",
                                "turquoise4"),
                    labels=c("30","40","50","60","70/80"))+
  scale_linetype_manual(labels=c("30","40","50","60","70/80"),
                        values = c(1,2,4,5,6))+
  labs(fill="Posted speed limit (km/h)",
       color="Posted speed limit (km/h)",
       lty="Posted speed limit (km/h)")

ggsave(plot = Spd_DistFor,
       "Spd_DistFor.png",
       width = 6, height = 5,
       units = "in",
       dpi=700)





```


# High speed land cover

Modelling showed that distance to forest and the amount of urban cover were more predictive of WVCs than the speed limit on the road, so let's compare the distribution of land cover types across the three higher speed limits that managers are more likely to be interested in.

```{r}

HighSpd <- subset(RoeDeer.df, speed>50)
HighSpdCollisions <- subset(HighSpd, response=="1")



ggplot(data = HighSpdCollisions,
       aes(dist_cover))+
  geom_histogram(aes(fill=f.speed),
                 alpha=0.7,
                 binwidth = sd(HighSpdCollisions$dist_cover))+
  facet_wrap(~f.speed)+
  scale_fill_viridis(discrete=TRUE,option="B", end=0.9)+
  scale_y_continuous(expand = c(0,0))+
  scale_x_continuous(expand = c(0,0))+
  theme(strip.background = element_rect(fill="white"))+
  xlab("Distance to forest cover class (m)")+
  labs(fill="Nearest posted speed limit (km/h)")


ggplot(data = subset(HighSpdCollisions, f.speed=="60"))+
  geom_histogram(aes(cover_fo),
                     fill="#C4aa23",
                 color="#C4aa23",
                 alpha=0.3,
                 bins=100)+
  geom_histogram(aes(cover_agri),
                     fill="#2c7c94",
                 color="#2c7c94",
                 alpha=0.3,
                 bins=100)+
   geom_histogram(aes(cover_urb),
                     fill="#a65852",
                 color="#a65852",
                 alpha=0.3,
                 bins=100)+
  scale_y_continuous(expand = c(0,0))+
  scale_x_continuous(expand = c(0,0))
  

```



## reformat data to be used in pie charts of cover types

For each speed, I need to put all the cover classes together in a single column with another column with the cover class and the speed category.

So I may be inefficient, but I first take each speed and create dummy variables to hold the cover categories

```{r}

###### 60 KMH

KMH60 <- subset(HighSpdCollisions, speed=="60")

COVERCat60 <-c(rep("FOR",nrow(KMH60)),
               rep("AGR",nrow(KMH60)),
               rep("WAT",nrow(KMH60)),
               rep("URB",nrow(KMH60))
)

COVER60 <- c(KMH60$cover_fo,
             KMH60$cover_agri,
             KMH60$cover_wa,
             KMH60$cover_urb)

SPD60 <- rep("60", length(COVER60))


###### 70 KMH

KMH70 <- subset(HighSpdCollisions, speed=="70")

COVERCat70 <-c(rep("FOR",nrow(KMH70)),
               rep("AGR",nrow(KMH70)),
               rep("WAT",nrow(KMH70)),
               rep("URB",nrow(KMH70))
)

COVER70 <- c(KMH70$cover_fo,
             KMH70$cover_agri,
             KMH70$cover_wa,
             KMH70$cover_urb)

SPD70 <- rep("70", length(COVER70))

###### 80 KMH



COVERCat80 <-c(rep("FOR",nrow(KMH80)),
               rep("AGR",nrow(KMH80)),
               rep("WAT",nrow(KMH80)),
               rep("URB",nrow(KMH80))
)
COVER80 <- c(KMH80$cover_fo,
             KMH80$cover_agri,
             KMH80$cover_wa,
             KMH80$cover_urb)

SPD80 <- rep("80", length(COVER80))


HighSpeedCover <- data.frame(
  cbind(
    Speed = c(SPD60,SPD70,SPD80),
    CoverCat = c(COVERCat60,COVERCat70,COVERCat80),
    CoverProp =c(COVER60,COVER70,COVER80)
  )
)

HighSpeedCover$CoverProp <- as.numeric(HighSpeedCover$CoverProp)

```


```{r}

ggplot(data = subset(HighSpeedCover, CoverProp>0),
       aes(CoverProp))+
  geom_histogram(aes(fill=CoverCat))+
  facet_wrap(~Speed)

library(wesanderson)
pal <-  wes_palette("Rushmore1",
                    type = "discrete")

HighSpdCover <- ggplot(data = subset(HighSpeedCover, CoverProp>0),
       aes(CoverProp))+
  geom_histogram(aes(fill=Speed),
                 alpha=0.7)+
  facet_wrap(~CoverCat)+
  scale_y_continuous(expand = c(0,0))+
  scale_x_continuous(expand=c(0,0))+
  scale_fill_manual(values=c("turquoise3","tan4","limegreen"))
  
  
ggsave(plot=HighSpdCover,
       "HighSpdCover.png",
       width = 6, height = 10,
       units = "in",
       dpi=700)



```



## Pie chart

```{r}


ggplot(data = HighSpeedCover, aes(CoverCat,CoverProp))+
  geom_col(aes(fill=CoverCat))+
  coord_polar("y", start = 0)+
  facet_wrap(~Speed)
  



```

