---
title: "Roe Deer Logistic Regression Analysis"
author: "Ashlee Mikkelsen"
date: "`r Sys.Date()`"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Introduction

This is the logistic regression file for the Roe deer- vehicle collision analysis for Viken kommune. Andreas, Rick and I took a contract from the kommune to estimate whether decreasing the speed limit on roads by 10 km/hr would reduce the number of vehicle collisions with Roe deer. 

# Environment

```{r}

rm(list=ls())

```

## Set working directory

```{r}

setwd("~Rprojects\RoeDeerVehCol")

```


## Load packages

```{r}

library(ggplot2)
library(lme4)
library(wiqid)
library(AICcmodavg)


```


## Load Data

```{r}

RoeDeer.df <- read.csv("Data\\df_VikenRoeDeer.csv",
                         header=TRUE)
RoeDeer.df$f.speed <- as.factor(RoeDeer.df$speed)

```

The original data file has 21090 observations and was processed by Rick prior to analysis. Explanation of data file:
Columns
1: Object ID- Unique identifyer for each record
2: id- secondary identifyer identifies the response category of the observation and the observation number within the response category (1_1 is the first observation of the 1 response category)
3: response- category 1= a collision 0= randomly drawn background points that are psuedo-absence points.
4: X- continuous variable UTM northing of the record
5: Y- continuous variable UTM easting of the record
6: speed- discrete numeric Speed limit of road segment closest straight line distance from the record
7: ADT-  (?)
8: efficiency-  continuous variable (0-1) indexing the straightness of the road in which 0= perfectly straigh and records greater than 0 have some level curvature. 
9: cover_fo- continuous variable (0-1) proportion of forest cover calculated as
10: cover_agri- continuous variable (0-1) proportion of forest cover calculated as
11: cover_wa- continuous variable (0-1) proportion of water calculated as
12: cover_urban- continuous variable (0-1) proportion of urban land cover calculated as
13: dist_cover- continuous variable staraight line distance from the record to (forest?) cover, calculated as, units= meters?
14: dist-agri- continuous variable staraight line distance from the record to agriculture land cover, calculated as, units= meters?
15: dist_water- continuous variable staraight line distance from the record to (forest?) cover, calculated as, units= meters?
16: dist_urban- continuous variable staraight line distance from the record to (forest?) cover, calculated as, units= meters?
17: geometry-
18: f.speed- categorical variable for speed


## Standardize continuos variables

```{r}



```


# Models

## Intercept only

```{r}
Models<-list()

Models[[1]]<-INT<-glm(response~1,
                        data = RoeDeer.df,
                        family = binomial(link="logit"))
summary(INT)

```


## Speed

### continuous discrete

```{r}
Models[[2]]<-SPD.c<-glm(response ~ speed,
                        data = RoeDeer.df,
                        family = binomial(link="logit"))
summary(SPD.c)

```

### categorical (factor)

```{r}

Models[[3]]<-SPD.f<-glm(response ~ f.speed,
                        data = RoeDeer.df,
                        family = binomial(link="logit"))
summary(SPD.f)

spd.c_coef <- (exp(0.02)/(1+exp(0.02)))


```



# Generate model table for comparison

```{r}

Modnames <- c("INTERCEPT", "SPEED_cont", "SPPED_fac")

##generate AICc table

aictab(cand.set = Models,
       modnames = Modnames, sort = TRUE, second.ord=TRUE)

##round to 4 digits after decimal point and give log-likelihood
print(aictab(cand.set = Models,
             modnames = Modnames,
             sort = TRUE),
      digits = 4,
      LL = TRUE)

write.csv(aictab(cand.set = Models,
                 modnames = Modnames,
                 sort = TRUE,
                 second.ord=TRUE),
          file="~Results\\ModelSelectionTable.csv")


```

